---
title: "Lesson 2 Intro To Spatial In R"
author: "James Gill"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
source("setup.R")
```

#Import colorado counties with tigris

```{r}
counties <- counties(state = "co")
```

#Import roads for Larimer county

```{r}
roads <- roads(state = "CO", county = "Larimer")
```

#Set tmap mode to interactive

```{r}
tmap_mode("view")
```

#Using qtm

```{r}
qtm(counties)+
  qtm(roads)
```

#Using tm_shape

```{r}
tm_shape(counties)+
  tm_polygons()+
tm_shape(roads)+
  tm_lines()
```

#Class counties

```{r}
class(counties)
```

#Filter Highway

```{r}
poudre_hwy <- roads %>% 
  filter(FULLNAME == "Poudre Canyon Hwy")

qtm(poudre_hwy)
```

#Point Data

```{r}
poudre_points <- data.frame(name = c("Mishawaka", "Rustic", "Blue Lake Trailhead"),
                            long = c(-105.35634, -105.58159, -105.85563),
                            lat = c(40.68752, 40.69687, 40.57960))
```

#Convert to spatial

```{r}
poudre_points_sf <- st_as_sf(poudre_points, coords = c("long", "lat"), crs = 4326)

qtm(poudre_hwy)+
  qtm(poudre_points_sf)
```

#Coordinate reference systems

```{r}
# see the CRS in the header metadata:
counties

#return just the CRS (more detailed)
st_crs(counties)
```

#CRS check

```{r}
st_crs(counties) == st_crs(poudre_points_sf)
```

#Transform layers

```{r}
poudre_points_prj <- st_transform(poudre_points_sf, st_crs(counties))
```

#Now check that they match

```{r}
st_crs(poudre_points_prj) == st_crs(counties)
```

#Elevation raster data

```{r}
elevation <- get_elev_raster(counties, z = 7)
```

```{r}
qtm(elevation)
```

```{r}
tm_shape(elevation)+
  tm_raster(style = "cont", title = "Elevation (m)")
```

#If we try this, we get an error

```{r}
elevation_crop <- crop(elevation, ext(poudre_hwy))
```

#Converting raster data

```{r}
elevation <- rast(elevation)
names(elevation) <- "Elevation"
```

#Terra functions

```{r}
elevation_crop <- crop(elevation, ext(roads))

tm_shape(elevation_crop)+
  tm_raster(style = "cont")
```

#Final map

```{r}
tm_shape(elevation, bbox = st_bbox(poudre_hwy))+
  tm_raster(style = "cont", title = "Elevation (m)")+
tm_shape(poudre_hwy)+
  tm_lines()+
tm_shape(poudre_points_prj)+
  tm_dots(size = 0.2)
```

#Save data

```{r}
write_sf(poudre_hwy, "data/poudre_hwy.shp")

write_sf(poudre_points_prj, "data/poudre_points.shp")
```

#Save data with terra

```{r}
writeRaster(elevation_crop, "data/elevation_larimer.tif")
```

#Save .RData files

```{r}
save(poudre_hwy, poudre_points_prj, file = "data/spatial_objects.RData")
```

#Remove .RData files

```{r}
rm(poudre_hwy, poudre_points_prj)
```

#Load .RData files

```{r}
load("data/spatial_objects.RData")
```

#Unpack .RDS

```{r}
saveRDS(elevation_crop, "data/elevation_crop.RDS")
```

```{r}
readRDS("data/elevation_crop.RDS") %>% rast()
```

#Read spatial file

```{r}
read_sf("data/poudre_hwy.shp")
```

#Read raster file

```{r}
rast("data/elevation_larimer.tif")
```

#Exercises

##1

```{r}
filter(counties, counties$NAME=="Larimer", counties$NAME=="Denver", counties$NAME=="Pueblo")
```

##2

```{r}
tm_shape(counties)+
    tm_polygons()
tm_shape(counties)+
  tm_fill(col="ALAND")
```

```{r}
tm_shape(counties)+
    tm_polygons()
tm_shape(counties)+
  tm_fill(col="AWATER")
```

##3

```{r}
poudre_elevation <- extract(elevation, poudre_points_prj)
ggplot(poudre_elevation)+
  geom_bar(mapping=aes(x=name, y=elevation))
```

##4
There are four features because we have four rows of data. 